Minimizing Time Required For Graduation

### 2010-12-13 paper/paper.tex
normalize sentence spacing; edit abstract
add determining dependents section, header, removed section 6

%%% HEADERS & FOOTERS
\usepackage{fancyhdr} % Set AFTER setting up the page geometry
\pagestyle{fancy} % options: empty , plain , fancy
\renewcommand{\headrulewidth}{0pt} % customise the layout...
\lhead{}\chead{}\rhead{}
\lfoot{}\cfoot{\thepage}\rfoot{}

@@

\abstract{The Bachelor's Degree takes four years to complete;
however, students often fail to complete it in this time frame. One
factor contributing to this is poor planning when registering for classes.
We present an algorithm to assist students in constructing a schedule 
that works towards completing a degree in the shortest amount of time 
possible. The algorithm presented uses heuristics to attempt to pick the best
class set each quarter, working towards a complete schedule quarterly by 
quarter.}
\pagebreak

@@

\subsubsection{Heuristic Value Function} To pick the classes taken in a given quarter,
we use a heuristic function which assigns
each class a value. When constructing the schedule, we pick classes with the highest heuristic value.
Recall that our goal is to
minimize the total number of quarters taken (and implicitly, maximize the total
number of courses taken per quarter). Therefore, our heuristic function gives higher
value to class that maximize the number of classes taken in a given quarter and in
future quarters.  One component to the heuristic function then must be the
number of dependent courses a given course offering has. If a course satisfies
prerequisites, then there are courses that we cannot
complete without it. Therefore, we want our heuristic function to favor it.

@@

@@

\section{Computational Results} In this section we will outline results obtained
by running computer simulations of schedule picking. We will computationally
examine what parameters in our heuristic function produce the most desirable
schedules.

>>>>

%%% HEADERS & FOOTERS
\usepackage{fancyhdr} % Set AFTER setting up the page geometry
\pagestyle{fancy} % options: empty , plain , fancy
\renewcommand{\headrulewidth}{0pt} % customise the layout...  
\lhead{Minimizing Time Required For Graduation}\chead{}\rhead{} \lfoot{}\cfoot{\thepage}\rfoot{}

@@

\abstract{The Bachelor's Degree takes four years to complete;
however, students often fail to complete it in this time frame.  One
factor contributing to this is poor planning when registering for classes.
We present an algorithm to assist students in constructing a schedule 
that works towards completing a degree in the shortest amount of time 
possible.  The algorithm presented uses heuristics to pick the best
class set each quarter, working towards a complete schedule quarter by
quarter.}
\pagebreak

@@

\subsubsection{Heuristic Value Function} To pick the classes taken in a given quarter,
we use a heuristic function which assigns
each class a value.  When constructing the schedule, we pick classes with the
highest heuristic value.  Recall that our goal is to minimize the total number
of quarters taken (and implicitly, maximize the total number of courses taken
per quarter).  Therefore, our heuristic function gives higher value to classes
that maximize both the number of classes taken in a given quarter and in future
quarters.  One component to the heuristic function then must be the number of
dependent courses a given course offering has.  If a course satisfies
prerequisites, then there are courses that we cannot complete without it.
Therefore, we want our heuristic function to favor it.

@@

\subsubsection{Determining $Dependents(c)$} We briefly address how to determine
the dependent value of a particular course. Figure~\ref{prereq} indicates the
method for determining the dependent value of a course; generate the graph of
prerequisites and then count the number of children in the subtree rooted at
a particular course node. If this was the only consideration though, we could
potentially assign a very high value to a course that is not related to
a student's major becuase that course had a large number of dependents. The
purpose of our system is to provide schedules that get a student through a major
as quickly as possible, so we should only value courses that lead to graduation.
The set $S$ describes the classes a student will need to graduate.

To assign values to courses, we start with the courses in $S$ and work backwards
using the prerequisite rules in $R$ to determine how to traverse the
prerequisite DAG. Traversing the graph using this method, we determine the value
of the course by counting the children in trees rooted at the current node in
the traversal. By starting with the courses in $S$ we ensure that we only given
values to courses that we need to take, instead of any courses with a large
number of dependents.

@@
